---
title: "Creating Tables - LaTeX Tables"
header-includes:
   - \usepackage{xfrac}
   - \floatplacement{table}{H}
format: 
  pdf:
    highlight: tango
    latex_engine: xelatex
    fig_width: 6
    fig_height: 6
mainfont: "Sabon"
sansfont: "Helvetica Neue UltraLight"
monofont: Inconsolata
always_allow_html: yes
---

```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false

# Load libraries
library(corrr)
library(knitr)
library(kableExtra)
library(tidyverse)
library(tidyr)

# Ignore NAs and replace with blank space
options(knitr.kable.NA = '')

# Import data
usnews = read_csv(file = "https://raw.githubusercontent.com/zief0002/musings/refs/heads/main/data/ed-schools-2018.csv") |> drop_na()
riverview = read_csv(file = "https://raw.githubusercontent.com/zief0002/musings/refs/heads/main/data/riverview.csv")

```


```{r}
#| label: tbl-01
#| tbl-cap: "Means and standard deviations for five measures of graduate programs of education."
#| echo: false


# Set up data frame
data.frame(
  Measure = c("Peer rating", "Acceptance rate for Ph.D. students", "Enrollment", "GRE score (verbal)", "GRE score (quantitative)"),
  M  = c(mean(usnews$peer), mean(usnews$doc_accept), mean(usnews$enroll), 
         mean(usnews$gre_verbal), mean(usnews$gre_quant)),
  SD = c(sd(usnews$peer), sd(usnews$doc_accept), sd(usnews$enroll), 
         sd(usnews$gre_verbal), sd(usnews$gre_quant))
) |>
  kable(
    col.names = c("Measure", "\\textit{M}", "\\textit{SD}"),
    align = c("l", "c", "c"),
    digits = 1,
    format = "latex",
    booktabs = TRUE,
    escape = FALSE
    )
```


\vspace{1in}


```{r}
#| label: tbl-02
#| tbl-cap: "Means and standard deviations for three measures of Riverview employees conditioned on sex."
#| echo: false

# Set up data frame
data.frame(
  Measure = c("Education level (in years)", "Seniority (in years)", "Income (in U.S. dollars"),
  M_f  = c(
    mean(riverview$education[riverview$gender == "female"]),
    mean(riverview$seniority[riverview$gender == "female"]),
    mean(riverview$income[riverview$gender == "female"])
  ),
  SD_f = c(
    sd(riverview$education[riverview$gender == "female"]),
    sd(riverview$seniority[riverview$gender == "female"]),
    sd(riverview$income[riverview$gender == "female"])
  ),
  M_m  = c(
    mean(riverview$education[riverview$gender == "male"]),
    mean(riverview$seniority[riverview$gender == "male"]),
    mean(riverview$income[riverview$gender == "male"])
  ),
  SD_m = c(
    sd(riverview$education[riverview$gender == "male"]),
    sd(riverview$seniority[riverview$gender == "male"]),
    sd(riverview$income[riverview$gender == "male"])
  )
) |>
  kable(
    col.names = c("Measure", "\\textit{M}", "\\textit{SD}", "\\textit{M}", "\\textit{SD}"),
    align = c("l", "c", "c", "c", "c"),
    digits = 0,
    format = "latex",
    booktabs = TRUE,
    escape = FALSE
  ) |>
  add_header_above(
    header = c(" " = 1, "Females" = 2, "Males" = 2)
  )
```


\newpage


```{r}
#| label: tbl-03
#| tbl-cap: "Means and confidence intervals (CIs) for three measures of Riverview employees conditioned on sex."
#| echo: false

# Set up data frame
data.frame(
  Measure = c("Education level (in years)", "\\hspace{2em} Female", "\\hspace{2em} Non-female", "Seniority (in years)", "\\hspace{2em} Female", "\\hspace{2em} Non-female", "Income (in thousand of U.S. dollars)", "\\hspace{2em} Female", "\\hspace{2em} Non-female"),
  M = c(NA, 16, 16, NA, 14, 16, NA, 48.9, 59.9),
  S = c(NA, 4, 5, NA, 7, 7, NA, 13.2, 14.2),
  LL = c(NA, 13.7, 13.5, NA, 10.7, 11.5, NA, 42.3, 51.7),
  UL = c(NA, 17.8, 19.1, NA, 17.5, 19.9, NA, 55.5, 68.1)
) |>
  kable(
    col.names = c("Measure", "\\textit{M}", "\\textit{SD}", "\\textit{LL}", "\\textit{UL}"),
    align = c("l", "c", "c", "c", "c"),
    format = "latex",
    booktabs = TRUE,
    escape = FALSE
  ) |>
  row_spec(row = 0, align = "c") |>
  add_header_above(c(" " = 3, "95% CI" = 2))
```


\vspace{1in}


```{r}
#| label: tbl-04
#| tbl-cap: "Intercorrelations for five measures of graduate programs of education."
#| echo: false
#| message: false

tab_04 = usnews |>
  select(peer, doc_accept, enroll, gre_verbal, gre_quant) |>
  correlate() |>
  shave(upper = TRUE) |>
  fashion(decimals = 2, na_print = "")  |>
  data.frame()

tab_04[1, 2] = "---"
tab_04[2, 3] = "---"
tab_04[3, 4] = "---"
tab_04[4, 5] = "---"
tab_04[5, 6] = "---"

tab_04 |>
  mutate(
    term = c("1. Peer rating", "2. Acceptance rate for Ph.D. students", "3. Enrollment",
                "4. GRE score (verbal)", "5. GRE score (quantitative)")
  ) |>
  kable(
    col.names = c("Measure", "1", "2", "3", "4", "5"),
    align = c("l", "c", "c", "c", "c", "c"),
    format = "latex",
    booktabs = TRUE,
    escape = FALSE
  )
```


\newpage


```{r}
#| label: tbl-05
#| tbl-cap: "Means, standard deivations, and intercorrelations for five measures of graduate programs of education."
#| echo: false
#| message: false

tab_05_2 = usnews |>
  select(peer, doc_accept, enroll, gre_verbal, gre_quant) |>
  correlate() |>
  shave(upper = TRUE) |>
  fashion(decimals = 2, na_print = "—")

tab_05_1 = data.frame(
  M = c(3.3, 40.1, 970, 154.9, 151.0),
  SD = c(0.5, 20.2, 665, 3.7, 4.4)
)

tab_05 = cbind(tab_05_2[, 1], tab_05_1, tab_05_2[, 2:6])

tab_05[1, 5:8] = ""
tab_05[2, 6:8] = ""
tab_05[3, 7:8] = ""
tab_05[4, 8] = ""

tab_05 |>
  mutate(
    `tab_05_2[, 1]` = c("1. Peer rating", "2. Acceptance rate for Ph.D. students", "3. Enrollment", "4. GRE score (verbal)", "5. GRE score (quantitative)")
  ) |>
  kable(
    col.names = c("Measure", "\\textit{M}", "\\textit{SD}", "1", "2", "3", "4", "5"),
    align = c("l", "c", "c", "c", "c", "c", "c", "c"),
    format = "latex",
    booktabs = TRUE,
    escape = FALSE
  ) 
```


\vspace{1in}


```{r}
#| label: tbl-06
#| tbl-cap: "Unstandardized coefficients for an OLS regression model fitted to estimate variation in peer ratings."
#| echo: false

tab_06 = lm(peer ~ 1 + doc_accept + enroll + gre_verbal + gre_quant, data = usnews) |>
  broom::tidy() |>
  mutate(
    term = c("Constant", "Acceptance rate for Ph.D. students", "Enrollment", 
             "GRE score (verbal)", "GRE score (quantitative)")
  )

tab_06 = tab_06[c(2:5, 1), ]

tab_06 |>
  kable(
    col.names = c("Predictor", "\\textit{B}", "\\textit{SE}", "\\textit{t}", "\\textit{p}"),
    align = c("l", "c", "c", "c", "c"),
    digits = c(NA, 2, 2, 2, 3),
    format = "latex",
    booktabs = TRUE,
    escape = FALSE
  )
```


\newpage


```{r}
#| label: tbl-07
#| tbl-cap: "Unstandardized coefficients and confidence intervals for an OLS regression model fitted to estimate variation in peer ratings."
#| echo: false

tab_07 = lm(peer ~ 1 + doc_accept + enroll + gre_verbal + gre_quant, data = usnews) |>
  broom::tidy(conf.int = TRUE) |>
  mutate(
    term = c("Constant", "Acceptance rate for Ph.D. students", "Enrollment", 
             "GRE score (verbal)", "GRE score (quantitative)")
  )

tab_07[c(2:5, 1), ] |>
  select(term, estimate, std.error, conf.low, conf.high)  |>
  kable(
    col.names = c("Predictor", "\\textit{B}", "\\textit{SE}", "\\textit{LL}", "\\textit{UL}"),
    align = c("l", "c", "c", "c", "c"),
    digits = c(NA, 2, 2, 2, 2),
    format = "latex",
    booktabs = TRUE,
    escape = FALSE
  ) |>
  row_spec(row = 0, align = "c") |>
  add_header_above(c(" " = 3, "95% CI" = 2))
```



\vspace{1in}


```{r}
#| label: tbl-08
#| tbl-cap: "Unstandardized and standardized coefficients afor an OLS regression model fitted to estimate variation in peer ratings."
#| echo: false

tab_08 = broom::tidy(lm(peer ~ 1 + doc_accept + enroll + gre_verbal + gre_quant, data = usnews))

tab_08[c(2:5, 1), ] |>
  mutate(
    term = c("Acceptance rate for Ph.D. students", "Enrollment", "GRE score (verbal)", 
             "GRE score (quantitative)", "Constant"),
    beta = c("-0.41", "0.69", "0.007", "0.32", "---"),
    p.value = scales::pvalue(p.value, accuracy = .001)
  ) |>
  select(term, estimate, beta, statistic, p.value) |>
  kable(
    col.names = c("Predictor", "\\textit{B}", "$\\beta$", "\\textit{t}", "\\textit{p}"),
    align = c("l", "c", "c", "c", "c"),
    digits = c(NA, 2, 2, 2, 3),
    format = "latex",
    booktabs = TRUE,
    escape = FALSE
  )
```




\newpage


```{r}
#| echo: false
#| message: false
#| results: "asis"
#| eval: false

library(stargazer)

lm.1 = lm(peer ~ 1 + gre_verbal + gre_quant, data = usnews)
lm.2 = lm(peer ~ 1 + doc_accept + enroll, data = usnews)
lm.3 = lm(peer ~ 1 + doc_accept + enroll + gre_verbal + gre_quant, data = usnews)

stargazer(lm.1, lm.2, lm.3,
          ci = TRUE,
          column.labels = c("Model 1", "Model 2", "Model 3"),
          covariate.labels = c("GRE score (verbal)", "GRE score (quantitative)", "Acceptance rate for Ph.D. students", "Enrollment"),
          dep.var.caption = "Outcome variable: Peer ratings",
          dep.var.labels = NULL,
          dep.var.labels.include = FALSE,
          type = "latex",
          keep.stat = c("rsq","ser"),
          star.cutoffs = NA,
          header = FALSE,
          table.placement = 'H',
          title = "Unstandardized Coefficients and Confidence Intervals for a Series of Regression Models Fitted to Data from $n=129$ Graduate Schools of Education to Predict Variation in Peer Ratings",
          omit.table.layout = "n"
)
```

\begin{table}[H]
\caption{Unstandardized coefficients and confidence intervals for a series of OLS regression models fitted to estimate variation in peer ratings.}
\label{}
\begin{tabular}{@{\extracolsep{5pt}}lccc}
\\[-1.8ex]\hline\\[1ex]
& Model 1 & Model 2 & Model 3 \\
\hline \\[-1.8ex]
GRE score (verbal) & 0.011 &  & 0.001 \\
& ($-$0.024, 0.046) &  & ($-$0.031, 0.033) \\
& & & \\
GRE score (quantitative) & 0.047 &  & 0.036 \\
& (0.017, 0.076) &  & (0.009, 0.063) \\
& & & \\
Acceptance rate for Ph.D. students &  & $-$0.013 & $-$0.010 \\
&  & ($-$0.017, $-$0.009) & ($-$0.014, $-$0.006) \\
& & & \\
Enrollment &  & 0.0001 & 0.0001 \\
&  & ($-$0.00004, 0.0002) & ($-$0.0001, 0.0002) \\
& & & \\
Constant & $-$5.488 & 3.769 & $-$1.857 \\
& ($-$8.683, $-$2.294) & (3.572, 3.967) & ($-$5.054, 1.340) \\
& & & \\
\hline \\[-1.8ex]
R$^{2}$ & 0.243 & 0.300 & 0.390 \\
RMSE & 0.429 & 0.413 & 0.389 \\
\hline
\end{tabular}
\end{table}


\newpage

```{r}
#| echo: false
#| message: false
#| results: "asis"
#| eval: false

stargazer(lm.1, lm.2, lm.3,
          column.labels = c("Model 1", "Model 2", "Model 3"),
          covariate.labels = c("GRE score (verbal)", "GRE score (quantitative)", "Acceptance rate for Ph.D. students", "Enrollment"),
          dep.var.caption = "Outcome variable: Peer ratings",
          dep.var.labels = NULL,
          dep.var.labels.include = FALSE,
          type = "latex",
          keep.stat = c("rsq","ser"),
          report = c("vcsp"),
          star.cutoffs = NA,
          header = FALSE,
          table.placement = 'H',
          title = "Unstandardized Coefficients and Confidence Intervals for a Series of Regression Models Fitted to Data from $n=129$ Graduate Schools of Education to Predict Variation in Peer Ratings",
          omit.table.layout = "n"
)
```



\begin{table}[H]
\caption{Unstandardized coefficients (standard errors) and $p$-Values for a series of OLS regression models fitted to estimate variation in peer ratings.}
\label{}
\begin{tabular}{@{\extracolsep{5pt}}lccc}
\\[-1.8ex]\hline\\[1ex]
& Model 1 & Model 2 & Model 3 \\
\hline \\[-1.8ex]
GRE score (verbal) & 0.011 &  & 0.001 \\
& (0.018) &  & (0.016) \\
& $p = 0.531$ &  & $p = 0.950$ \\
& & & \\
GRE score (quantitative) & 0.047 &  & 0.036 \\
& (0.015) &  & (0.014) \\
& $p = 0.003$ &  & $p = 0.012$ \\
& & & \\
Acceptance rate for Ph.D. students &  & $-$0.013 & $-$0.010 \\
&  & (0.002) & (0.002) \\
&  & $p = 0.000$ & $p = 0.00000$ \\
& & & \\
Enrollment &  & 0.0001 & 0.0001 \\
&  & (0.0001) & (0.0001) \\
&  & $p = 0.239$ & $p = 0.347$ \\
& & & \\
Constant & $-$5.488 & 3.769 & $-$1.857 \\
& (1.630) & (0.101) & (1.631) \\
& $p = 0.002$ & $p = 0.000$ & $p = 0.258$ \\
& & & \\
\hline \\[-1.8ex]
R$^{2}$ & 0.243 & 0.300 & 0.390 \\
RMSE & 0.429 & 0.413 & 0.389 \\
\hline
\hline \\[-1.8ex]
\end{tabular}
\end{table}


\newpage

\begin{table}[H]
\caption{Unstandardized Coefficients and Confidence Intervals for a Series of Regression Models Fitted to Data from $n=129$ Graduate Schools of Education to Predict Variation in Peer Ratings}
\label{}
\begin{tabular}{@{\extracolsep{5pt}}lccc}
\\[-1.8ex]\hline\\[1ex]
& Model 1 & Model 2 & Model 3 \\
\hline \\[-1.8ex]
GRE score (verbal) & 0.011 (0.018) &  & 0.001 (0.016) \\
& $p = 0.531$ &  & $p = 0.950$ \\
& & & \\
GRE score (quantitative) & 0.047 (0.015) &  & 0.036 (0.014) \\
& $p = 0.003$ &  & $p = 0.012$ \\
& & & \\
Acceptance rate for Ph.D. students &  & $-$0.013 (0.002) & $-$0.010 (0.002) \\
&  & $p = 0.000$ & $p = 0.00000$ \\
& & & \\
Enrollment &  & 0.0001 (0.0001) & 0.0001 (0.0001)\\
&  & $p = 0.239$ & $p = 0.347$ \\
& & & \\
Constant & $-$5.488 (1.630) & 3.769 (0.101) & $-$1.857 (1.631) \\
& $p = 0.002$ & $p = 0.000$ & $p = 0.258$ \\
& & & \\
\hline \\[-1.8ex]
R$^{2}$ & 0.243 & 0.300 & 0.390 \\
RMSE & 0.429 & 0.413 & 0.389 \\
\hline
\hline \\[-1.8ex]
\end{tabular}
\end{table}




